# Dimensionality Reduction

```{r}
source(here::here("R", "dimReducFunction.R"))
```

```{r}
num_data <- College[sapply(College, is.numeric)]
corr_matrix <- cor(num_data)
```

<div class="hscroll-plot">
<div class='plot-container'>
```{r}
pairs(College[sapply(College, is.numeric)], gap=0, pch=16, cex=0.4)
```
</div>

<div class='plot-container'>
```{r}
corrplot.mixed(
  corr_matrix,
  lower = 'number',
  upper = 'pie',
  order = 'AOE',
  tl.cex = 0.7,        # riduce la dimensione delle etichette
  tl.col = "black",    # colore del testo
  tl.pos = "lt",       # posizione delle etichette (left-top)
  number.cex = 0.7     # grandezza dei numeri
)
```
</div>
</div>

### PCA {-}

```{r}
pca <- prcomp(num_data, scale. = TRUE)
```

<div class="hscroll-plot">
<div class='plot-container'>
```{r}
screePlot(pca)
```
</div>

<div class='plot-container'>
```{r}
screePlot(pca, TRUE)
```
</div>
</div>

<div class="hscroll-plot">
<div class='plot-container'>
```{r}
fviz_pca_var(pca,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )
```
</div>

<div class='plot-container'>
```{r}
fviz_pca_ind(pca,
             col.ind = "cos2", # Color by the quality of representation
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             label = 'none'
             )
```
</div>

<div class='plot-container'>
```{r}
groups <- as.factor(College$Private)

# Plot con fviz_pca_ind()
fviz_pca_ind(pca,
             col.ind = groups, # Colora in base a 'Private'
             palette = c("#00AFBB", "#FC4E07"), # Puoi cambiarli
             addEllipses = TRUE,
             ellipse.type = "confidence",
             legend.title = "Private",
             label = 'none')
```
</div>
</div>

**PCA Loadings (|loading| > 0.3)**

<div class="hscroll-plot" style="height: 700px;">
<div class='plot-container'>
```{r}
n_pc <- 17
threshold <- 0.3

plotLoadingsTable(pca, n_pc, threshold)
```
</div>
</div>

#### Choosing the number of PCs {-}

Let's apply **Kaiser's rule**

```{r fig.width=12, fig.height=5, echo=FALSE, message=FALSE, warning=FALSE}
kaiserRule(pca)
```

The PC four is on the line and the variable that it explain better are inside the first PC.

Then we select only the first 3 components:

* **PC1 – Academic Prestige and Student Spending**: This component captures the level of academic prestige and quality of a college. High values indicate selective institutions with good resources and qualified faculty.
* **PC2 – Size and Enrollment Volume**: This component appears to reflect the size of the institution and the scale of its admissions process. Low values correspond to large, highly attended colleges.
* **PC3 – Personal and Book Expenses**: Personal and Book Expenses: A component that seems to reflect student-related out-of-pocket expenses. Negative values indicate higher costs for personal living and study materials.
