---
site: bookdown::bookdown_site
documentclass: book
output:
  bookdown::gitbook:
    css: ../docs/style.css
    toc_depth: 2
    number_sections: true
---

# Index

_Author_: [Daniele Tambone](https://www.linkedin.com/in/daniele-tambone-b5733616a/)

_GitHub project_: [College-Analysis](https://github.com/DanieleT25/College-Analysis)

_Date_: `r Sys.Date()`


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  fig.width = 8,   # larghezza in pollici
  fig.height = 5   # altezza in pollici
)
```

```{r, include=FALSE}
library(ISLR2)
library(here)
library(DT)
library(skimr)
library(DataExplorer)
#library(dplyr)
#library(plotly)
library(gamlss)
#library(gamlss.mx)
library(mixR)
```


## Table of Contents {-}

- [Abstract](abstract.html)
- [1. Introduction](introduction.html)
- [2. Univariate Analysis](univariate-analysis.html)
- [3. Dimensionality Reduction](dimensionality-reduction.html)
- [4. Cluster Analysis](clustering.html)
- [5. Conclusion](conclusion.html)

This is the report of Unsupervised learning.
