### Cluster validation

#### External clustering validation
This is possible apply only on the result of K-means++ ($k=2$), because it is the unique clustering algorithem that identify the beahiour of the private and public college.

```{r}
true_labels_num <- as.numeric(College$Private)

clust_stats <- cluster.stats(
  d = dist(scale(College[, -1])), 
  true_labels_num,
  pred_clustersKmeans
)
```

```{r}
results_table <- data.frame(
  Metric = c("Corrected Rand Index", "Meila's VI Index"),
  Value = c(clust_stats$corrected.rand, clust_stats$vi)
)

kable(results_table, digits = 4, align = "c") %>%
  kable_styling(full_width = TRUE)
```

Meila’s VI Index is not very high (its range is $(0, \infty)$), but the Corrected Rand Index is low.

#### Stability measures

```{r}
# Hierarchical - Euclidean - Average - 2 clusters
stab_hcEuclAvg2 <- get_stability_metrics_custom(
  df = df,
  original_cluster = hcEuclAvg2$cluster,
  get_cluster_del = function(df_del) {
    cutree(hclust(dist(df_del, method = "euclidean"), method = "average"), k = 2)
  },
  method = "euclidean"
)

# Hierarchical - Euclidean - Average - 5 clusters
stab_hcEuclAvg5 <- get_stability_metrics_custom(
  df = df,
  original_cluster = hcEuclAvg5$cluster,
  get_cluster_del = function(df_del) {
    cutree(hclust(dist(df_del, method = "euclidean"), method = "average"), k = 5)
  },
  method = "euclidean"
)

# Hierarchical - Manhattan - Average - 2 clusters
stab_hcManhAvg2 <- get_stability_metrics_custom(
  df = df,
  original_cluster = hcManhAvg2$cluster,
  get_cluster_del = function(df_del) {
    cutree(hclust(dist(df_del, method = "manhattan"), method = "average"), k = 2)
  },
  method = "manhattan"
)

# Hierarchical - Manhattan - Average - 8 clusters
stab_hcManhAvg8 <- get_stability_metrics_custom(
  df = df,
  original_cluster = hcManhAvg8$cluster,
  get_cluster_del = function(df_del) {
    cutree(hclust(dist(df_del, method = "manhattan"), method = "average"), k = 8)
  },
  method = "manhattan"
)

# KMeans - 3 clusters
stab_kmeans3 <- get_stability_metrics_custom(
  df = df,
  original_cluster = kmeans3$cluster,
  get_cluster_del = function(df_del) {
    eclust(df_del, "kmeans", k = 3, graph = FALSE)$cluster
  },
  method = "euclidean"
)

# KMeans++ - 2 clusters
stab_kmeansPlus2 <- get_stability_metrics_custom(
  df = df,
  original_cluster = kmeansPlus2$cluster,
  get_cluster_del = function(df_del) {
    KMeans_rcpp(df_del, 2)$cluster
  },
  method = "euclidean"
)

# KMedoids - Euclidean - 3 clusters
stab_kmedoidsEucl3 <- get_stability_metrics_custom(
  df = df,
  original_cluster = kmedoidsEucl3$clustering,
  get_cluster_del = function(df_del) {
    pam(df_del, 3, metric = "euclidean")$clustering
  },
  method = "euclidean"
)

# KMedoids - Manhattan - 3 clusters
stab_kmedoidsManh3 <- get_stability_metrics_custom(
  df = df,
  original_cluster = kmedoidsManh3$clustering,
  get_cluster_del = function(df_del) {
    pam(df_del, 3, metric = "manhattan")$clustering
  },
  method = "manhattan"
)
```

```{r}
stability_table <- tibble::tibble(
  Method = c(
    "HC - Euclidean - Avg - k=2",
    "HC - Euclidean - Avg - k=5",
    "HC - Manhattan - Avg - k=2",
    "HC - Manhattan - Avg - k=8",
    "KMeans - k=3",
    "KMeans++ - k=2",
    "KMedoids - Euclidean - k=3",
    "KMedoids - Manhattan - k=3"
  ),
  APN = c(
    stab_hcEuclAvg2$APN,
    stab_hcEuclAvg5$APN,
    stab_hcManhAvg2$APN,
    stab_hcManhAvg8$APN,
    stab_kmeans3$APN,
    stab_kmeansPlus2$APN,
    stab_kmedoidsEucl3$APN,
    stab_kmedoidsManh3$APN
  ),
  AD = c(
    stab_hcEuclAvg2$AD,
    stab_hcEuclAvg5$AD,
    stab_hcManhAvg2$AD,
    stab_hcManhAvg8$AD,
    stab_kmeans3$AD,
    stab_kmeansPlus2$AD,
    stab_kmedoidsEucl3$AD,
    stab_kmedoidsManh3$AD
  ),
  ADM = c(
    stab_hcEuclAvg2$ADM,
    stab_hcEuclAvg5$ADM,
    stab_hcManhAvg2$ADM,
    stab_hcManhAvg8$ADM,
    stab_kmeans3$ADM,
    stab_kmeansPlus2$ADM,
    stab_kmedoidsEucl3$ADM,
    stab_kmedoidsManh3$ADM
  ),
  FOM = c(
    stab_hcEuclAvg2$FOM,
    stab_hcEuclAvg5$FOM,
    stab_hcManhAvg2$FOM,
    stab_hcManhAvg8$FOM,
    stab_kmeans3$FOM,
    stab_kmeansPlus2$FOM,
    stab_kmedoidsEucl3$FOM,
    stab_kmedoidsManh3$FOM
  )
)

kable(stability_table, digits = 4, align = "c", booktabs = TRUE) %>%
  kable_styling(full_width = TRUE, bootstrap_options = c("striped", "hover"))
```

```{r}
get_best_methods <- function(metrics_df) {
  best_methods <- list()

  best_methods[["Best APN"]] <- metrics_df$Method[which.min(metrics_df$APN)]
  best_methods[["Best AD"]]  <- metrics_df$Method[which.min(metrics_df$AD)]
  best_methods[["Best ADM"]] <- metrics_df$Method[which.min(metrics_df$ADM)]
  best_methods[["Best FOM"]] <- metrics_df$Method[which.min(metrics_df$FOM)]

  for (name in names(best_methods)) {
    cat(name, ": \"", best_methods[[name]], "\"\n", sep = "")
  }
}
```

```{r}
get_best_methods(stability_table)
```

From these stability measures, we can see that the most stable clustering algorithms are HC with Euclidean distance and average linkage ($k=2$), and K-means ($k=3$). However, the first solution yields only two clusters (with one of them has only one element). Therefore, K-means with $k=3$ is preferable.

Finally, we should choose between k-means with $k=3$ and k-means++ with $k=2$.
