### Model Based Clustering

One disadvantage of traditional clustering methods, such as hierarchical and partitioning clustering algorithms, is that they are largely heuristic and not based on formal models.

An alternative is model-based clustering that use **probabilistic assumptions**.

```{r}
invisible(
    capture.output(
      mbc <- Mclust(dataScale)
    )
)
```

```{r}
fviz_mclust(mbc, "BIC", palette = "jco")
```

```{r}
summary(mbc$BIC)
```

Using the **Mclust** function, the best model selected is **VVE**, which corresponds to clusters with variable volume and shape but equal orientation, with $k=9$.

```{r}
summary(mbc)
```

However, from a graphical point of view, the result is very difficult to interpret.

<div class="hscroll-plot">
<div class='plot-container'>
```{r}
fviz_mclust(mbc, "classification", palette = "jco", geom = "point", ellipse.type = "norm", ggtheme = theme_minimal(), labelsize = 0)
```
</div>

<div class='plot-container'>
```{r}
fviz_mclust(mbc, "uncertainty", palette = "jco")
```
</div>
</div>
